Word Identification Under Multimodal Uncertainty

### 2017-02-02 write-up/main.tex
draft new

Identifying the visual referent of a spoken word -- that a particular insect is referred to by the word ``bee'' -- requires both the ability to process and integrate multi-modal input and the ability to identify categories under uncertainty. How do these tasks interact with one another? We introduce a task that allows us to examine how adults identify word categories under joint uncertainty in the auditory and visual modalities. We propose a Bayesian model of the task which provides an optimal baseline. Model predictions are tested in two experiments where category recognition is made under two kinds of uncertainty: ambiguity and noise. In both cases, the optimal baseline model explains much of the variance in human judgments. But when one modality had noise added to it, human perceivers systematically preferred the unperturbed modality to a greater extent than the optimal baseline model did.

@@

 % (e.g., the visual modality in the case of concrete objects)
Language uses symbols expressed in one modality (e.g., the auditory modality, in the case of speech) to communicate about the world, which we perceive through many different sensory modalities. Consider hearing someone yell ``bee!" at a picnic, as a honeybee buzzes around the food. Determining reference involves processing the auditory information and linking it with other perceptual signals (e.g., the visual image of the bee, the sound of its wings, the sensation of the bee flying by your arm). But processing this linguistic input also requires categorization. On the linguistic side,  individual sounds must be identified as members of phonetic categories for the word to be understood. On the referential side, individual sensory impressions must be categorized as coming from a bee. Once these two sets of category-identification tasks have been completed, then a listener can make the referential link between word and world.

This comprehension task takes place in a noisy world. On the auditory side, individual acoustic word tokens are almost always ambiguous with respect to the particular sequence of phonemes they represent \cite<e.g.,>{hillenbrand1995}. And some tokens may be distorted additionally by mispronunciation or ambient noise. Perhaps the speaker was yelling ``pea" and not ``bee.'' Similarly, a sensory impression may not be enough to make a definitive identification of a visual category.\footnote{In the general case, language can of course be visual as well as auditory, and object identification can be done through many modalities. For simplicity, we focus on audio-visual matching here.} Perhaps the quickly flying insect was a beetle or a fly instead.

Thus, establishing reference requires reasoning under a great deal of uncertainty. Work on phoneme categorization has suggested that perceivers encode tokens probabilistically: they base their decisions on these probabilities and adjust their behavior when these probabilities change \cite<e.g.,>{clayard08}. However, previous work in this line of research has largely focused on the unimodal case, that is, when category uncertainty is only in one modality \cite<e.g.,>{feldman2009}. A few studies have explored some aspects of cross-modal processing  \cite<e.g.,>{bejjanki2011, kleinschmidt2015}. This work has focused on the specific case of phoneme recognition from speech and lip movement, however, where information from both modalities is tightly correlated.

In the present work, we study the case of word recognition, where a word is an \textit{arbitrary} mapping between a sound category and a visual category. We test participants on their ability to process audio-visual stimuli and use them to recognize the underlying word category. Crucially, this recognition is made under uncertainty about both the auditory category (a single phoneme, in our case) and the visual category (the concept being referred to). Our goal is to explore how adults deal with such multimodal ambiguity in a categorization task. To this end, we propose a model that combines uncertainty in an optimal way, and we use it as a baseline for interpreting human data. More precisely, we examine the extent to which human data conform to, or deviate from the baseline, and whether the deviation involves a systematic preference for a particular modality.

The paper is organized as follows. First, we introduce our audio-visual word recognition task. We next present the model and explain how it provides an optimal baseline. Then we present two behavioral experiments where we test category recognition under audio-visual uncertainty. In Experiment 1, audio-visual tokens are clear but their category membership is ambiguous. In Experiment 2, we intervene by adding additional perceptual noise to one modality. While participants show no modality preference in Experiment 1, in Experiment 2 they over-rely on visual input when the auditory modality is noisy.

@@

We introduce a new task that tests audio-visual category recognition. We use two visual categories (cat and dog) and two auditory categories (/b/ and /d/ embedded in the minimal pair /aba/-/ada/). For each participant, an arbitrary pairing is set between the auditory and the visual categories, leading to two audio-visual word categories (e.g., dog-/aba/, cat-/ada/).

In each trial, participants are presented with an audio-visual target (the prototype of the target category), immediately followed by an audio-visual test stimulus (Figure \ref{fig:task}). The test stimulus may differ from the target in both the auditory and the visual components. The auditory component is sampled from a continuum along the second formant (F2) linking the words /aba/ and /ada/. The visual component is sampled from a morph that links the prototypical pictures of the dog and the cat.  After these two presentations, participants are instructed to press ``same'' or ``different.''

This task is similar to the task introduced by \citeA{sloutsky2003} and used in subsequent research to probe audio-visual encoding. However, unlike this previous line of research, here participants are not asked whether the two audio-visual presentations are identical. Instead, the task is category-based: They are asked to press `same' if they think the second item (the test) belonged to the same category as the first (e.g.,  dog-/aba/), even if there is a slight difference in the word, in the object, or in both. They are instructed to press `different' only if they think that the second stimulus was an instance of the other word category (cat-/ada/).

The task also included trials where pictures were hidden (audio-only) or where sounds were muted (visual-only). These unimodal trials provide us with participants' categorization functions for the auditory and visual categories and are used as inputs to the optimal model, described below.

@@

The basis of our ideal observer model is that individual categorization functions from each modality should be combined optimally.
In each modality, we have two categories: /ada/ ($A=1$) and /aba/ ($A=2$) in the auditory dimension, and \textit{cat} ($V=1$) and \textit{dog} ($V=2$) in the visual dimension.
% \footnote{The identity of the categories was randomized in the experimental setup.}
We assume, for the sake of simplicity, that the probability of membership in each category is normally distributed:

@@

In the bimodal condition, participants see word tokens with audio-visual input, and have to make a categorization decision. We represent word tokens as vectors in the audio-visual space, $\mathbf{w}=(a,v)$.
A word category $W$ is defined as the joint distribution of auditory and visual categories. It can be characterized with a bivariate normal distribution in the audio-visual space:
$$ p(\mathbf{w} | W) \sim  N(M_W, \Sigma_W) $$

@@

We have two word categories: dog-/aba/ ($W_1$) and cat-/ada/ ($W_2$). Participants can be understood as choosing one of these two word categories (figure \ref{fig:space}). For an ideal ``recognizer'', the probability of choosing category 2 when presented with an audio-visual instance $\mathbf{w}=(a,v)$ is the posterior probability of this category:

@@

In sum, the posterior \ref{eq:pred} provides a baseline for how probabilities that characterize uncertainty in each modality can be combined to make categorical decision about bimodal input.

@@

In experiment 1, we test the predictions of the model in the case where uncertainty is due to perceptual ambiguity in terms of category membership. This experiment had two goals. First, we wanted to examine the extent to which human responses corresponded to the predictions of the ideal observer model. Second, if there was a deviation from the baseline, we wanted to be able to measure whether this deviation involved a systematic preference for a particular modality.

@@

For  auditory stimuli, we used the continuum introduced in \citeA{vroomen2004}, a 9-point /aba/--/ada/ speech continuum created by varying the frequency of the second (F2) formant in equal steps. We selected 5 equally spaced points from the original continuum by keeping the end-points (prototypes) 1 and 9, as well as points 3, 5, and 7 along the continuum. For visual stimuli, we used a morph continuum introduced in \citeA{freedman2001}: a continuous set of images was generated from three cat prototypes and three dog prototypes. From the original 14 points, we selected 5 points as follows: we kept the item that seemed most ambiguous (point 8), the 2 preceding points (i.e., 7 and 6) and the 2 following points (i.e., 8 and 9). The 6 and 9 points along the morph were quite distinguishable, and we took them to be our prototypes.

@@

We told participants that an alien was naming two objects: a dog, called /aba/ in the alien language, and a cat, called /ada/. In each trial, we presented the first object (the target) on the left side of the screen simultaneously with the corresponding sound. The target was always the same (e.g., dog-/aba/). The second sound-object pair (the test) followed on the other side of the screen after 500ms and varied in its category membership. For both the target and the test, visual stimuli were present for the duration of hte sound clip ($\sim$800ms). We instructed participants to press ``S'' for same if they thought the alien was naming another dog-/aba/, and ``D'' for different if they thought the alien was naming a cat-/ada/. For each participant, we randomized the sound-object mapping as well as the identity of the target.

@@

\subsection{Results}
%
% \begin{figure}[h]
% \centering
% \includegraphics[width=3.25in\textwidth]{correlation.png}
% \caption{Human responses vs. the baseline predictions. Error bars are 95\% confidence intervals. In experiment 1, the baseline and the fit explained 89\%, and 94\% of total variance, respectively. In experiment 2, the baseline and the fit explained 91\%, and 97\% of total variance, respectively}
% \label{fig:corr}
% \end{figure}

\begin{figure*}[t]
\centering
\includegraphics[width=.83\textwidth]{Exp.png}
\caption{Proportion of ``different'' judgments as a function of auditory distance. Solid lines represent average human responses (left), predictions of the baseline (middle), and the bimodal fit (left). Dashed lines represent average human responses in the unimodal conditions. Colors represent values in the visual continuum.}
\label{fig:Exp}
% \vspace{10ex}
\end{figure*}

% Let's take the auditory-only case as an example,
\subsubsection{Unimodal conditions}

Average categorization judgments and fits are shown in Figure \ref{fig:unimodal}. Auditory judgements were somewhat sharper and more pronounced than visual judgments. For an ideal recognizer, the probability of choosing category 2 (that is, to answer ``different'') when presented with e.g., an audio instance $a$, is the posterior probability of this category $p(A_2|a)$. If we assume that both categories have equal variances, the posterior probability reduces to:

@@

For this model (as well all other models), we fixed the values of the means to be the end-points of the corresponding continuum: $\mu_{A1}=0$ and $\mu_{A2}=4$ (and similarly $\mu_{V1}=0$, and $\mu_{V2}=4$). To determine the values of the bias and the variance, we fit a model for each modality, collapsed across participants. For the auditory modality, we got $\epsilon_A=-0.20$ and $\sigma^2_A=2.04$. For the visual modality, we got $\epsilon_V=-0.11$ and $\sigma^2_V=3.34$.
% In figure \ref{fig:unimodal} (top) we plotted responses in the unimodal conditions as well as the unimodal fits.

We fit a model to human responses in the bimodal condition, collapsed across participants, finding $\epsilon=-0.32$, $\sigma^2_{Ab}=5.00$ and $\sigma^2_{Vb}=7.27$.

\subsubsection{Baseline model}

We derived the predictions of the baseline model by using the values of the variances derived from the unimodal conditions, and the response bias derived from the bimodal condition, and by substituting these values into the expression of the posterior in Eq.~\ref{eq:pred}. Figure \ref{fig:Exp} shows participants' responses in the bimodal condition (left), as well as the prediction of the baseline (middle), and the bimodal fit models (right).
% Correlation of human data and the bimodal baseline/fitted model are shown in Figure \ref{fig:corr} (top).

@@

We found negative values for in all response bias terms, which suggests a general bias toward answering ``different.''  This bias is probably due to the categorical nature of the task: when two items are ambiguous but perceptually different, this could cause a slight preference for ``different'' over ``same.''

@@

We next analyzed whether there was a preference for one or the other modality when making decisions in the bimodal condition, beyond that explained by the variance in categories implied by the unimodal responses. This preference would manifest as a deviation from the decision threshold predicted by the baseline model. The decision threshold is defined as the set of values in the audio-visual space along which the posterior (Eq.~\ref{eq:pred}) is equal to 0.5. The decision threshold takes the following form:
% As illustrated in \ref{fig:space}, above this line participants are more likely to choose one word category, and below this line, they are more likely to choose the other word category.

@@

If the slope derived from the fit is bigger than the slope of the baseline, this finding would suggest a general preference for the auditory modality (similarly, a smaller slope would suggest a preference for the visual modality). The limit cases are when there is exclusive reliance on the auditory cue (a vertical line), and where there is exclusive reliance on the visual (a horizontal line). Figure \ref{fig:pref} (top left) shows the decision threshold in the audio-visual space with a constant intercept; the fit to human data  (solid black line) was very close to the baseline threshold (red line). Non-parametric resampling sampling of the data showed no evidence of a deviation from the baseline (\ref{fig:pref}, bottom left).

@@

% Compare human responses (figure \ref{fig:Exp}, top left) to the baseline (figure \ref{fig:Exp}, top middle).
Qualitatively, participants' judgments were similar to the predictions of the optimal baseline model (when fit to the unimodal data). Consider the contrast between the auditory-only case (dashed black line in Figure \ref{fig:Exp}, top left) and the bimodal case (solid colored lines).  Higher certainty in the visual modality generally influenced responses, with greater visual distance leading to more ``different'' ratings and less visual distance leading to more ``same'' ratings. Indeed, we found that the baseline model explained almost 90\% of the variance in judgments ($r^2=0.89$). In addition, as described above, we did not see any evidence for a modality preference.

But although we see a qualitative resemblance data and baseline model, there were quantitative differences. Model predictions were more influenced by the visual modality at the auditory midpoint (the point of highest uncertainty) than human judgements, and were more compressed at the endpoints (the points of lowest uncertainty). Formally speaking, this trend translated into an increase in the value of the variance associated with each modality (e.g., compare the variance values in the baseline model to the values derived from the bimodal fit).  Since the variance characterizes the reliability of a given cue, it seems as if bimodal presentation caused reliance on both modalities to decrease. This effect could possibly be attributed to the arbitrary nature of the form-meaning mapping. In fact, previous studies suggest that while redundant multi-modal information improves performance (e.g., determining the frequency of a bouncing ball from visual and auditory cues), arbitrary mappings generally tends to hinder performance \citeA<for review, see>{robinson2010}.

@@

In Experiment 1, we tested the ability of people to recognize a word category under ambiguous tokens. In particular, the tokens were clear but their category membership was uncertain. In real life situations, however, even the identity of the token can be uncertain due to various distorting factors. In order to account for this kind of uncertainty in the auditory modality, we can imagine \cite<following>{feldman2009}, that the speaker generates a target production $t$ from an auditory category
$t | A \sim N(\mu_{A}, \sigma^2_{A})$. The listener, in a noisy condition, does not hear exactly this produced target, but the target perturbed by normally distributed noise: $a | t \sim N(t, \sigma^2_{N})$. When we integrate over $t$, we get:

@@

In Experiment 2, we explored the effect of noise on performance in our task. We tested a case where one modality was ambiguous and noisy (auditory), and where the other modality was ambiguous but not noisy (visual). We were interested to know if participants would treat this new source of uncertainty as predicted by the optimal baseline, and whether noise in one modality would lead to some systematic preference for the non-noisy modality.

@@

We used the same visual stimuli as in Experiment 1. We also used the same auditory stimuli , but we convolved each item with Brown noise of amplitude 1 using the audio editor Audacity (2.1.2). The procedure was exactly the same as in the previous experiment, except that target stimuli were presented with the new noisy auditory stimuli.

@@

\subsubsection{Baseline model}

We generated the predictions of the baseline model by using the values of the variances derived from the unimodal conditions, and the response bias derived from the bimodal condition, and by plugging these values into Eq.~\ref{eq:pred}.
Results are shown in Figure \ref{fig:Exp} (bottom).
 % Correlation of human data to the bimodal baseline and the fit are shown in figure \ref{fig:corr} (bottom).

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.4\textwidth]{preference.png}
\caption{Top: decision thresholds in the audio-visual space. Red dotted line is the prediction of the baseline. Blue dotted lines are cases where modality preference is twice as strong as the baseline. Solid line is the threshold derived from human data. Bottom: comparison of the threshold slope between the baseline and the fit to human data. Error bars are 95\% confidence intervals computed through non-parametric bootstrap sampling (10.000 samples)}
\label{fig:pref}
\end{figure}

We found, similar to experiment 1, that participants generally showed qualitative patterns similar to the baseline model ($r^2 = .91$). But we also found a similar discrepancy at the quantitative level: There was a decrease in the reliability of both modalities, as could be inferred from the increase in the variance associated with the auditory modality and the visual modality in the bimodal fit. An interesting difference with Experiment 1, however, was that participants' decision threshold suggested a preference for the visual modality (the non-noisy modality). Indeed non-parametric bootstrap sampling over the data showed there to be a significant decrease in the value of the slope.

\section{Conclusion}

In this work, we tested word category recognition under audio-visual uncertainty. We modeled uncertainty as a probability distribution over tokens from auditory and visual continua. Our baseline model instantiated an ideal observer, optimally combining participant data from unimodal decisions. The model's predictions were examined in two experiments where we tested two kinds of uncertainty: in experiment 1, audio-visual tokens were clear but their category membership was ambiguous. In experiment 2, input to one modality was ambiguous, and input to the other was ambiguous as well as noisy. In both experiments, the ideal observer model explained the majority of variance in participant judgments. In both experiments, however, we found that bimodal presentation caused the reliability of both modalities to decrease relative to the ideal observer.

We explored whether part of the unexplained variance could involve to a systematic preference for a given modality. Our analysis showed no evidence of such modality preference in experiment 1. However, we found a preference for the non-noisy modality in experiment 2.

This paper addressed the problem of recognizing words whose underlying phonemic and semantic categories were supposed to be familiar. An intriguing challenge is to understand how these underlying categories are acquired in the first place. Previous developmental work addressed cases when one modality was supposed to provide a stable and certain cue to the other modality \cite<e.g.,>{waxman1995, yeung09}. The present paper invites investigations on category acquisition in the more general case where there is uncertainty in both modalities.

>>>>

Identifying the visual referent of a spoken word -- that a particular insect is referred to by the word ``bee'' -- requires both the ability to process and integrate multi-modal input and the ability to reason under uncertainty. How do these tasks interact with one another? We introduce a task that allows us to examine how adults identify words under joint uncertainty in the auditory and visual modalities. We propose a ideal observer model of the task which provides an optimal baseline. Model predictions are tested in two experiments where word recognition is made under two kinds of uncertainty: category ambiguity and distorting noise. In both cases, the ideal observer model explains much of the variance in human judgments. But when one modality had noise added to it, human perceivers systematically preferred the unperturbed modality to a greater extent than the ideal observer model did.

@@

Language uses symbols expressed in one modality (e.g., the auditory modality, in the case of speech) to communicate about the world, which we perceive through many different sensory modalities. Consider hearing someone yell ``bee!" at a picnic, as a honeybee buzzes around the food. Determining reference involves processing the auditory information and linking it with other perceptual signals (e.g., the visual image of the bee, the sound of its wings, the sensation of the bee flying by your arm).

This multi-modal integration task takes place in a noisy world. On the auditory side, individual acoustic word tokens are almost always ambiguous with respect to the particular sequence of phonemes they represent, which is due to the inherent variability of how a phonetic category is realized acoustically \cite<e.g.,>{hillenbrand1995}. And some tokens may be distorted additionally by mispronunciation or ambient noise. Perhaps the speaker was yelling ``pea" and not ``bee.'' Similarly, a sensory impression may not be enough to make a definitive identification of a visual category.\footnote{In the general case, language can of course be visual as well as auditory, and object identification can be done through many modalities. For simplicity, we focus on audio-visual matching here.} Perhaps the quickly flying insect was a beetle or a fly instead.

Thus, establishing reference requires reasoning under a great deal of uncertainty in both modalities. The goal of this work is to characterize such reasoning.  Imagine, for example, that someone is uncertain whether they heard ``pea" or ``bee'', does it make them rely more on the visual modality (e.g., the object being pointed at)? Vice versa, if they are not sure if they saw a bee or a fly, does that make them rely more on the auditory modality (i.e., the label)? More importantly, when input in both modalities is uncertain to varying degrees, do they weight each modality according to its relative reliability, or do they over-rely on a particular modality?

In this paper, we propose a probabilistic framework where such reasoning can be expressed in a precise way. We characterize uncertainty, in each modality, with a probability distribution, and we predict ideal responses by combining these probabilities in a optimal way. Our work can be seen as an extension to previous Bayesian models of phoneme identification \cite<e.g.,>{feldman2009}, where, instead of a unimodal input, we model a bimodal one. A few studies have explored some aspects of audio-visual processing in a probabilistic framework  \cite<e.g.,>{bejjanki2011, kleinschmidt2015}. In these studies, the researchers focused on the specific case of phoneme recognition from speech and lip movement, however, where information from both modalities is tightly correlated.

In the present work, we study, rather, the case of arbitrary mapping between sounds and visual objects. We test participants on their ability to process audio-visual stimuli and use them to recognize the underlying word. More precisely we study the case where both the word's form and the word's referent are ambiguous, and we examine the extent to which humans conform to, or deviate from the predictions of the ideal observer model. Moreover, since some previous studies on audio-visual processing documented cases of modality preference (i.e., when people rely predominantly on the visual modality \cite<e.g., >{colavita74} or the auditory modality \cite<e.g., >{sloutsky2003}), we will  explore if, in our task as well, people show evidence of a modality preference.

The paper is organized as follows. First, we introduce our audio-visual recognition task. We next present the ideal observer model. Then we present two behavioral experiments where we test word recognition under audio-visual uncertainty. In Experiment 1, audio-visual tokens are ambiguous with respect to their category membership. In Experiment 2, we intervene by adding  noise to one modality. In both experiments participants show qualitative patterns of optimal behaviour. Moreover, while participants show no modality preference in Experiment 1, in Experiment 2 they over-rely on visual input when the auditory modality is noisy.

@@

We introduce a new task that tests audio-visual word recognition. We use two visual categories (cat and dog) and two auditory categories (/b/ and /d/ embedded in the minimal pair /aba/-/ada/). For each participant, an arbitrary pairing is set between the auditory and the visual categories, leading to two audio-visual word categories (e.g., dog-/aba/, cat-/ada/).

In each trial, participants are presented with an audio-visual target (the prototype of the target category), immediately followed by an audio-visual test stimulus (Figure \ref{fig:task}). The test stimulus may differ from the target in both the auditory and the visual components.  After these two presentations, participants are instructed to press ``same'' or ``different.''

This task is similar to the task introduced by \citeA{sloutsky2003} and used in subsequent research to probe audio-visual encoding. However, unlike this previous line of research, here participants are not asked whether the two audio-visual presentations are identical. Instead, the task is category-based: They are asked to press `same' if they think the second item (the test) belonged to the same category as the first (target) (e.g.,  dog-/aba/), even if there is a slight difference in the word, in the object, or in both. They are instructed to press `different' only if they think that the second stimulus was an instance of the other word category (cat-/ada/).

The task also included trials where pictures were hidden (audio-only) or where sounds were muted (visual-only). These unimodal trials provide us with participants' categorization functions for the auditory and visual categories and are used as inputs to the ideal observer model, described below.

@@

The basis of our ideal observer model is that individual categorization functions from each modality should be combined optimally.
In each modality, we have two categories: /ada/ ($A=1$) and /aba/ ($A=2$) in the auditory dimension, and \textit{cat} ($V=1$) and \textit{dog} ($V=2$) in the visual dimension.
We assume, for the sake of simplicity, that the probability of membership in each category is normally distributed:

@@

In the bimodal condition, participants see word tokens with audio-visual input, and have to make a categorization decision. We define word tokens as vectors in the audio-visual space, $\mathbf{w}=(a,v)$.
A word category $W$ is defined as the joint distribution of auditory and visual categories. It can be characterized with a bivariate normal distribution in the audio-visual space:
$$ p(\mathbf{w} | W) \sim  N(M_W, \Sigma_W) $$

@@

We have two word categories: dog-/aba/ ($W_1$) and cat-/ada/ ($W_2$). Participants can be understood as choosing one of these two word categories (Figure \ref{fig:space}). For an ideal observer, the probability of choosing category 2 when presented with an audio-visual instance $\mathbf{w}=(a,v)$ is the posterior probability of this category:

@@

In sum, the posterior \ref{eq:pred} provides the predictions of ideal observer for how probabilities that characterize uncertainty in each modality can be combined to make categorical decision about bimodal input.

@@

In Experiment 1, we test the predictions of the model in the case where uncertainty is due to similar auditory categories, and similar visual categories. Crucially, the similarity is such that the distributions overlap. To simulate such uncertainty in a controlled fashion, we use a continuum along the second formant (F2) linking the words /aba/ and /ada/, and we use a morph that links the prototypical pictures of the dog and the cat.

@@

For  auditory stimuli, we used the continuum introduced in \citeA{vroomen2004}, a 9-point /aba/--/ada/ speech continuum created by varying the frequency of the second (F2) formant in equal steps. We selected 5 equally spaced points from the original continuum by keeping the end-points (prototypes) 1 and 9, as well as points 3, 5, and 7 along the continuum. For visual stimuli, we used a morph continuum introduced in \citeA{freedman2001}: a continuous set of images was generated from three cat prototypes and three dog prototypes. From the original 14 points, we selected 5 points as follows: we kept the item that seemed most ambiguous (point 8), the 2 preceding points (i.e., 7 and 6) and the 2 following points (i.e., 9 and 10). The 6 and 10 points along the morph were quite distinguishable, and we took them to be our prototypes.

@@

We told participants that an alien was naming two objects: a dog, called /aba/ in the alien language, and a cat, called /ada/. In each trial, we presented the first object (the target) on the left side of the screen simultaneously with the corresponding sound. The target was always the same (e.g., dog-/aba/). The second sound-object pair (the test) followed on the other side of the screen after 500ms and varied in its category membership. For both the target and the test, visual stimuli were present for the duration of the sound clip ($\sim$800ms). We instructed participants to press ``S'' for same if they thought the alien was naming another dog-/aba/, and ``D'' for different if they thought the alien was naming a cat-/ada/. For each participant, we randomized the sound-object mapping as well as the identity of the target.

@@

\subsection{Results}
%
% \begin{figure}[h]
% \centering
% \includegraphics[width=3.25in\textwidth]{correlation.png}
% \caption{Human responses vs. the ideal observer predictions. Error bars are 95\% confidence intervals. In experiment 1, the ideal observer and the fit explained 89\%, and 94\% of total variance, respectively. In experiment 2, the ideal observer and the fit explained 91\%, and 97\% of total variance, respectively}
% \label{fig:corr}
% \end{figure}

\begin{figure*}[t]
\centering
\includegraphics[width=.83\textwidth]{Exp.png}
\caption{Proportion of ``different'' judgments as a function of auditory distance. Solid lines represent average human responses (left), predictions of the ideal observer (middle), and the bimodal fit (left). Dashed lines represent average human responses in the unimodal conditions. Colors represent values in the visual continuum.}
\label{fig:Exp}
% \vspace{10ex}
\end{figure*}

% Let's take the auditory-only case as an example,
\subsubsection{Unimodal conditions} this is the case where the pictures were hidden, or where the sounds were muted.
Average categorization judgments and fits are shown in Figure \ref{fig:unimodal}. Auditory judgements were somewhat sharper and more pronounced than visual judgments. The fit was done as follows. For an ideal recognizer, the probability of choosing category 2 (that is, to answer ``different'') when presented with an audio instance $a$, is the posterior probability of this category $p(A_2|a)$. If we assume that both categories have equal variances, the posterior probability reduces to:

@@

For this model (as well all other models), we fixed the values of the means to be the end-points of the corresponding continuum: $\mu_{A1}=0$ and $\mu_{A2}=4$ (and similarly $\mu_{V1}=0$, and $\mu_{V2}=4$). To determine the values of the bias and the variance, we fit a model for each modality, collapsed across participants. For the auditory modality, we obtained $\epsilon_A=-0.20$ and $\sigma^2_A=2.04$. For the visual modality, we obtained $\epsilon_V=-0.11$ and $\sigma^2_V=3.34$.

\subsubsection{Bimodal condition}
We fit a model to human responses in the bimodal condition, collapsed across participants, finding $\epsilon=-0.32$, $\sigma^2_{Ab}=5.00$ and $\sigma^2_{Vb}=7.27$. The fit explained XX of the variance.

\subsubsection{Ideal observer model}

We derived the predictions of the ideal observer model by using the values of the variances derived from the unimodal conditions, and the response bias derived from the bimodal condition, and by substituting these values into the expression of the posterior in Eq.~\ref{eq:pred}. 
Figure \ref{fig:Exp} shows participants' responses in the bimodal condition (left), as well as the prediction of the ideal observer (middle), and the bimodal fit models (right).

@@

We found negative values in all response bias terms, which suggests a general bias toward answering ``different.''  This bias is probably due to the categorical nature of our same-different task: when two items are ambiguous but perceptually different, this could cause a slight preference for ``different'' over ``same''.

@@

We next analyzed whether there was a preference for one or the other modality when making decisions in the bimodal condition, beyond that explained by the variance in categories implied by the unimodal responses. This preference would manifest as a deviation from the decision threshold predicted by the ideal observer model. The decision threshold is defined as the set of values in the audio-visual space along which the posterior (Eq.~\ref{eq:pred}) is equal to 0.5. The decision threshold takes the following form:

@@

If the slope derived from the bimodal fit is greater than the slope of the ideal observer, this finding would suggest a general preference for the auditory modality (similarly, a smaller slope would suggest a preference for the visual modality). The limit cases are when there is exclusive reliance on the auditory cue (a vertical line), and where there is exclusive reliance on the visual (a horizontal line). Figure \ref{fig:pref} (top left) shows the decision threshold in the audio-visual space with a constant intercept; the fit to human data  (solid black line) was very close to the ideal observer threshold (red line). Non-parametric resampling of the data showed no evidence of a deviation from the slope of the ideal observer (\ref{fig:pref}, bottom left).

@@

Qualitatively, participants' judgments were similar to the predictions of the ideal observer model (remember that the latter was obtained by optimally combining fits to the unimodal data). Consider, for example, the contrast between the auditory-only case (dashed black line in Figure \ref{fig:Exp}, top left) and the bimodal case (solid colored lines).  Higher certainty in the visual modality generally influenced responses, with greater visual distance leading to more ``different'' ratings and less visual distance leading to more ``same'' ratings. Similar observations can be made about the contrast between the visual-only case and the bimodal case.  Overall, we found that the ideal observer model explained almost 90\% of the variance in judgments ($r^2=0.89$).

But although we see a qualitative resemblance between human data and the ideal observer model, there were quantitative differences. For example, model predictions were more influenced by the visual modality at the auditory midpoint (the point of highest uncertainty) than human judgements, and were more compressed at the endpoints (the points of lowest uncertainty).

Formally speaking, there was an increase in the value of the variance associated with each modality. More precisely, whereas the ideal observer model predicted the weights to be proportional to $1/\sigma^2_{A}$ and $1/\sigma^2_{V}$, for the auditory and the visual modalities, respectively (see expression \ref{eq:pred}), the fit to human data suggested that the real weights were proportional to $1/\sigma^2_{Ab}$ and $1/\sigma^2_{Vb}$. Interestingly, our analysis of modality preference showed that the relative values of these variances (i.e., $\frac{\sigma^2_V}{\sigma^2_A}$ vs. $\frac{\sigma^2_{Vb}}{\sigma^2_{Ab}}$) were almost the same (Figure \ref{fig:pref}, left). This means two things: 1) the bimodal presentation caused reliance on both modalities to decrease, and 2) this decrease did not affect the relative weighting of both modalities, i.e., participants were weighting modalities according to their relative reliability. The latter explains the qualitative resemblance between the predictions of the ideal observer and human data, and the former explains the quantitative discrepancy.
\newline

In Experiment 1, we tested the ability of people to recognize a a word under audio-visual uncertainty linked to category membership. We found that, generally speaking, participants followed the ideal observer model in that they weighted modalities according to their relative reliability. In real life situations, however, tokens can undergo distortions due to noisy factors in the environment. In Experiment 2, we try to account for this additional level of uncertainty.

@@

Imagine that the speaker generates a target production $t$ from an auditory category
$t | A \sim N(\mu_{A}, \sigma^2_{A})$. In Experiment 1, we assumed that the observed could directly retrieve this production token. But if the observer is in a noisy environment, they do not hear exactly this produced target, but the target perturbed by noise, which we assume, following \citeA{feldman2009}, that it is normally distributed: $a | t \sim N(t, \sigma^2_{N})$. When we integrate over $t$, we get:

@@

In this experiment, we explored the effect of this added noise\footnote{Note that we are considering environmental noise, which is different from the noise inherent to perception.} on performance in our task. We tested a case where one modality was ambiguous and noisy (auditory), and where the other modality was ambiguous but not noisy (visual). We were interested to know if participants would treat this new source of uncertainty as predicted by the ideal observer model, and whether noise in one modality would lead to some systematic preference for the non-noisy modality.

@@

We used the same visual stimuli as in Experiment 1. We also used the same auditory stimuli, but we convolved each item with Brown noise of amplitude 1 using the audio editor Audacity (2.1.2).  The signal-to-noise ratio was XXX dB. The procedure was exactly the same as in the previous experiment, except that test stimuli were presented with the new noisy auditory stimuli.

@@

\subsubsection{ideal observer model}

We generated the predictions of the ideal observer model by using the values of the variances derived from the unimodal conditions, and the response bias derived from the bimodal condition, and by substituting these values into the expression of the posterior in Eq.~\ref{eq:pred}.
Results are shown in Figure \ref{fig:Exp} (bottom).

Participants' decision threshold suggested a preference for the visual modality (the non-noisy modality). Indeed non-parametric resampling of the data showed there to be a decrease in the value of the slope (\ref{fig:pref}, right).

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.4\textwidth]{preference.png}
\caption{Top: decision thresholds in the audio-visual space. Red dotted line is the prediction of the ideal observer. Blue dotted lines are cases where modality preference is twice as strong as the ideal observer. Solid line is the threshold derived from human data. Bottom: comparison of the threshold slope between the ideal observer and the fit to human data. Error bars are 95\% confidence intervals computed through non-parametric bootstrap sampling (10.000 samples)}
\label{fig:pref}
\end{figure}

We found, similar to Experiment 1, that participants generally showed qualitative patterns similar to the ideal observer model ($r^2 = .91$). But we also found a similar discrepancy at the quantitative level, suggesting an effect of bimodal presentation on performance. Formally speaking, the ideal observer model predicted the modality weights to be proportional to $1/(\sigma^2_{A}+\sigma^2_{N})$ and $1/\sigma^2_{V}$, for the auditory and the visual modalities, respectively. The fit to human data suggested that the empirical weights were proportional to $1/\sigma^2_{Ab}$ and $1/\sigma^2_{Vb}$. An interesting difference with Experiment 1, however, was that participants had a clear preference for the non-noisy modality, as the values of the relative variances $\frac{\sigma^2_V}{\sigma^2_A + \sigma^2_N}$ vs. $\frac{\sigma^2_{Vb}}{\sigma^2_{Ab}}$ were different (\ref{fig:pref}, right). This preference affected the relative weighting, where, contrary to Experiment 1, the visual modality now has a greater weight than what could be expected from the relative reliability alone.

It is important to understand that this preference was not the mere consequence of the added noise increasing the variance of the auditory modality, since this increase was already accounted for in the ideal observer model. The preference was, rather, a form of over-reliance on the visual modality.

\section{General discussion}

Understanding language requires both the ability to integrate multi-modal input, and the ability to deal with uncertainty. In this work, we explored a case where both abilities were at play. We studied the case of identifying a word when both its form (auditory) and its referent (visual) were ambiguous with respect to their category membership (Experiment 1), and when the form was perturbed with additional noise (Experiment 2). We introduced a model that instantiated an ideal observer, predicting how information from each modality could be combined in an optimal way. In both experiments, participants showed the qualitative patterns of the ideal observer.

There were, however, quantitative differences: the audio-visual presentation caused the reliance on both modalities to decrease to a lower extent than the ideal observer model expected. This phenomenon could be attributed to the arbitrary nature of the form-meaning mapping. Previous studies suggest, indeed, that while redundant multi-modal information improves performance (e.g., determining the frequency of a bouncing ball from visual and auditory cues), arbitrary mappings generally tends to hinder performance \cite<for review, see>{robinson2010}. Another possibility could be that the ideal model failed to capture some other features of multi-modal integration. Further work is needed to understand this result.
